---
title: "Exploring the BRFSS data"
output: 
  html_document: 
    fig_height: 4
    highlight: pygments
    theme: spacelab
---

## Setup

### Load packages

```{r load-packages, message = FALSE}
library(ggplot2)
library(dplyr)
library(maps)
```

### Load data

```{r load-data}
load("brfss2013.RData")
```



* * *

## Part 1: Data

The observations in th BRFSS 2013 are collected via phone calls (both cell and 
landline). Landline interviewees are randomly selected adults within a household, 
cell interviewees reside in either private residences or college housing.

This makes the information only generalizable to individuals who have:
1. A phone of either type
2. Stable living situation

While this may represent a majority of the population, it excludes many groups 
who may engage in behavior qualifying as making them "at risk." Especially for the 
topic of `immunization` this makes the data non-generalizable to epidemiology studies (due to how herd 
immunity's effectiveness is susceptible to deviations from "normalcy" by small groups)

* * *

## Part 2: Research questions

**Research quesion 1:**
What is the relation between mental health and state residency?  
**Research quesion 2:**
Do Students who live on campus have different sleeping habits?  
**Research quesion 3:**
Do the financially insecure live in a harsher world? (subject to data availability)
snctmony and misphlpf

* * *

## Part 3: Exploratory data analysis

NOTE: Insert code chunks as needed by clicking on the "Insert a new code chunk" 
button (green button with orange arrow) above. Make sure that your code is visible
in the project you submit. Delete this note when before you submit your work.

**Research quesion 1:**

```{r}
sadstate <- brfss2013 %>%
  select(X_state, menthlth, addepev2) %>%
  filter(menthlth <= 31) %>% # Can't have more bad days than there are in a month
  na.omit()
colnames(sadstate) <- c("state", "mntl", "dd")

# We'll be subsetting this data, so make a function for plotting
plot.bad.days <- function(sadstate) {
  p <- ggplot(data = sadstate, aes(x = mntl)) +
    geom_bar() +
    labs(title = "Poor Mental Health Days per Respondent",
        x = "Number of Bad Days in the Past 30",
        y = "Respondents")
  return(p)
}

plot.bad.days(sadstate)
```

The data are heavily right skewed, so mean is unreliable. Let's subset further and look at only respondents who have had at least one bad day in the past month.

```{r}
sadstate <- sadstate %>%
  filter(mntl > 0)

plot.bad.days(sadstate)
```

Now we see a multimodal ditribution, that resembles a binomial distribution consisting of a right-skewed Gaussian at 2 and a narrow peak at 30.  
What happens if we strip out that peak at 30?

```{r}
plot.bad.days(sadstate %>% filter(mntl < 30)) +
  # https://stackoverflow.com/questions/49709563/highlight-every-nth-bar-using-geom-bar
  geom_bar(aes(fill = (mntl %% 5 == 0 | mntl == 2 | mntl == 7), group = mntl), position = "identity") +
  theme(legend.position = "none") +
  ggtitle("Poor Mental Health Days per Respondent, Common Responses Highlighted")
```

Taking into account the multimodality of the data (something similar to a "rounding bias" is occuring, 
respondees seemed to commonly reflect that they'd had a couple bad days or a bad week, otherwise they were more likely to go by multiples of 5), 
there is a clear right skew when one does *not* consider individuals for whom every day within the scope of their interview was one of poor mental health - 
one might suspect these people suffer from a depressive disorder, let's quickly check on that.

```{r}
dep_chk <- sadstate %>%
  group_by(dd)

dep_chk %>%
  summarise(mean(mntl), median(mntl))

ggplot(data = dep_chk %>% filter(mntl < 30), aes(x = dd, y = mntl / 30 * 100)) +
  geom_violin(aes(fill = dd),
              draw_quantiles = c(0.25, 0.5, 0.75)) +
  stat_summary(fun.y = mean, geom = "point", shape = 24) +
  labs(title = "Relationship of Depression and Bad Days, with Quantiles",
       x = "Diagnosed with Depression",
       y = "Percentage of Poor Mental Health Days in Past 30") +
  theme(legend.position = "None")

ggplot(data = dep_chk, aes(x = mntl)) +
  geom_bar(aes(fill = dd), position = position_dodge(width = 0.9)) +
  scale_y_continuous() +
  guides(fill = guide_legend(title = "Diagnosed Depression"))
```

People with diagnosed depression represent a majority of those who had 15 or more poor mental health days in the past 30 days.  

Unsurprisingly, people with diagnosed depressive disorders experience worse mental health!

```{r}
natl_avg_bad_days <- mean(sadstate$mntl)
natl_med_bad_days <- median(sadstate$mntl)

bad_days_by_state <- sadstate %>%
  group_by(state) %>%
  summarise(state_avg = mean(mntl), state_med = median(mntl)) %>%
  mutate(avg_dev = state_avg - natl_avg_bad_days,
         med_dev = state_med - natl_med_bad_days)

p <- ggplot(data = bad_days_by_state %>% arrange(desc(state_avg)),
            aes(x = reorder(state, state_avg), y = state_avg,
                labels = state_avg)) +
  geom_hline(yintercept = natl_avg_bad_days, color = "red") +
  geom_segment(aes(x = reorder(state, state_avg), y = natl_avg_bad_days,
                 xend = state, yend = state_avg)) +
  geom_point(stat = "identity") +
  theme(axis.text.x = element_text(angle=90)) +
  labs(title = "Average Poor Mental Health Days per State",
       x = "State",
       y = "Average Days per Month of Poor Mental Health") +
  coord_flip()
 

p

# TODO: Make a divergent color map centered at 0, use dev and map states with deviation
range_avg_dev <- range(bad_days_by_state$avg_dev)
range_med_dev <- range(bad_days_by_state$med_dev)
# geom_polygon(data = bad_days_by_state, aes(fill = dev)) +
# scale_fill_gradient(colors = c(blue, white, orange),
#                    breaks = c(-1, 0, 1))

usa <- map_data("usa")
```



**Research quesion 2:**

```{r}
students <- brfss2013 %>%
  filter(employ1 == 'A student') %>%
  select(sleptim1, cclghous) %>%
  na.omit()

nonstudents <- brfss2013 %>%
  filter(employ1 != 'A student') %>%
  select(sleptim1) %>%
  na.omit()

ggplot(data = nonstudents, aes(x = sleptim1)) +
  geom_bar()

ggplot(data = students %>%
         group_by(cclghous),
       aes(x = sleptim1)) +
  geom_bar(aes(fill = cclghous, group = cclghous), position = "identity")

students %>%
  group_by(cclghous) %>%
  summarise(mean(sleptim1), median(sleptim1))
```



**Research quesion 3:**
"Financial struggle"" is defined as having worried about rent or food in the year prior to being surveyed.

```{r}
# 

harsh_world <- brfss2013 %>%
  select(misphlpf, scntmony, scntmeal) %>%
  na.omit() %>%
  mutate(strgl = (scntmony != "Never" | scntmeal != "Never")) %>%
  group_by(strgl)

ggplot(harsh_world, aes(misphlpf, group = strgl)) +
  geom_bar(aes(y = ..prop..,
               fill = factor((strgl))),
           position = position_dodge()) +
  guides(fill = guide_legend(title = "Struggled in Past Year")) +
  labs(title = "Financial Struggles and Mental Illness Perception",
       x = "\"People Are Generally Caring Toward People With Mental Illness\"",
       y = "Proportion of Respondents") +
  coord_flip()
  

brkdn <- harsh_world %>% summarise(
  sa = mean(misphlpf == "Agree strongly"),
  a = mean(misphlpf == "Agree slightly"),
  na = mean(misphlpf == "Neither agree nor disagree"),
  d = mean(misphlpf == "Disagree slightly"),
  sd = mean(misphlpf == "Disagree strongly")
)
```

```{r}
#TODO: Analyze relationships between being diagnosed with depression, financial struggle, access to healthcare


```